### 2024-08-12 Slides/slides.Rmd
Update slides

@@

### Distribution of Economic Activity across the world

![Nightlight](Figures/The_earth_at_night_greyscale.png)

@@

---

# Introduction
.pull-left[
- In 1825 the Agger isthmus* broke by 1834 a natural channel had formed 
- 700 years prior a similar channel had silted up
- Does trade cause growth?
- Geography is more than soil quality. We are slaves of geomorphology.

#### What will be demonstrated:
  - The channel caused trade 
  - The channel caused a local population boom
  - Mechanism
  - Evidence for the reverse natural experiment in 1100s

.pull-right[
```{r echo=FALSE, out.width="100%", fig.align='center'}
knitr::include_graphics("Figures/Petersen 1877.png")
```
*Petersen (1877)*

???
**3/45**  
- Understanding what causes wealth (wellbeing, etc.) is at the heart of Economics 
- Understanding the interplay between geography (environment) and Economics is vital in the 21st century

@@

.pull-left[
### Institutions vs geography
- Institutions are the primary driver of economic development (Acemoglu et al., 2001)
- Geography acts through institutions (Rodrik et al., 2004)
- Geography plays a role (which is smaller than instituions) (Ketterer and Rodŕıguez-Pose, 2018)
- .red123[**Important to verify in an age of climate change**]
]

.pull-right[
### Geography
- Henderson et al. (2018): Geography explain 47 pct of nightlight variation
- Even large disruptions does not change long-run outcomes (Davis and Weinstein, 2002) $\rightarrow$ Path dependence or first nature geography?
- Bleakly and Lin (2012): Path dependence is part of the answer
- What about first nature? Extreme long-run: Diamond (1997); Allen et al. (2023); Matranga (2024)
- .green123[**'Human' time scale missing**]
]

@@

#### Pre-event: 
- 1085-1208: Natural western channel closed because of gradual land rises (since the last ice age)
- 1672: Eastern Limfjord town Aalborg 2nd largest in DK 
- Other Limfjord towns lacked far behind under Aalborg's local hedgemony

>>>>

```{css echo=FALSE}
.remark-slide-number {
  display: none;
}
```

@@

### Distribution of Economic Activity across the world
```{r echo=FALSE, out.width="75%", fig.align='center'}
knitr::include_graphics("Figures/The_earth_at_night_greyscale.png")
```

@@

.pull-left-wide[
# A Perfect Storm
- In 1825 a storm breached the narrow Agger Istmus
- In 1834 it became navigable
- The same event happened in reverse in the 1100s

- Time variation in first-nature geography: Difference-in-difference to study the causal effect of geography on economic development

- **Data:** Population, occupation, demography (fertility/internal migration), archaeological findings $\rightarrow$ **First nature geography determines the location of prosperity**
]

@@

.pull-left[
### Institutions vs geography
- Institutions are the primary driver of economic development (Acemoglu et al., 2001)
- Geography acts through institutions (Rodrik et al., 2004)
- Geography plays a role (which is smaller than instituions) (Ketterer and Rodríguez-Pose, 2018)
]

.pull-right[
### Geography
- Henderson et al. (2018): Geography explain 47 pct of nightlight variation
- Even large disruptions does not change long-run outcomes (Davis and Weinstein, 2002) $\rightarrow$ Path dependence or first nature geography?
- Bleakly and Lin (2012): Path dependence is part of the answer
- What about first nature? Extreme long-run: Diamond (1997); Allen et al. (2023); Matranga (2024)
]

@@

#### Pre-event: 
- 1085-1208: Natural western channel closed because of gradual land rises (since the last ice age)
- Limfjord towns lacked far behind the rest of the country
